Hospitalizations for psychiatric conditions among adolescents during the COVID-19 pandemic

### 2021-10-06 psy_peds_3.Rmd
adding percentage count, and its with percentage

### Percentage 
```{r}
thpd_psyPer <- count_icd %>%
  ggplot(aes(x = time_p, y = percentage_psy, fill=period)) +
  scale_fill_manual(values = cbPalette[c(3,4)]) +
  geom_vline(xintercept = as.Date(pandemic_start_date),
             linetype = "dashed") +
  labs(y = "Percentage",
       x = paste0("Date (by ", time_period,")"),
       title = paste0("Percentage of patients with mental health related ICD codes (per ", time_period,")")) + 
  geom_smooth(method = "lm", formula = y ~ splines::ns(x, 2), se = TRUE, aes(colour=period), colour= "black", size = 0.5)+
  geom_point()
thpd_psyPer
```
### Ratio: patients with psychiatric condition / patients without psychiatric condition

@@

ratios_patients_with_without_psy <- patients_without_psy_conditions_counts %>%
  full_join( patients_with_psy_conditions_counts, by=c("time_p", "period")) %>%
  mutate( ratio = count_psy_patients/count_no_psy_patients)

ratio_psy_non_psy <- ratios_patients_with_without_psy %>%
  ggplot(aes(x = time_p, y = ratio, fill = period, color = period)) +
  geom_point() +
  geom_vline(xintercept = as.Date(pandemic_start_date),
             linetype = "dashed") +
  scale_fill_manual(values = cbPalette) +
  scale_color_manual(values = cbPalette) +
  labs(y = "Ratio (with psy conditions/without psy conditions)",
       x = paste0("Date (by ", time_period,")"),
       title = paste0("Ratio patient with vs without psychiatric conditions ( per ", time_period,")"))

@@

@@

```{r}
# Model without seasonality
time_series_analysis_plotting <- function(model, df_count_psy, count_col, clearance_period = FALSE) {
  output_model <- broom::tidy(model)
  coeff_slope <- round(output_model[output_model$term == "periodduring_pandemic:time", "estimate"][[1]], 3)
  pvalue_slope <- round(output_model[output_model$term == "periodduring_pandemic:time", "p.value"][[1]], 3)
  annotation <- paste0("Lockdown slope: ", coeff_slope, " (pvalue: ", pvalue_slope, ")")
  pred_model <- predict(model, patient_count_monthly_psy, type = "response", interval = "confidence") %>% as.data.frame()
  pred_model_b <- predict(model, mutate(patient_count_monthly_psy, period = "before_pandemic"), type = "response", interval = "confidence") %>%
    as.data.frame()
  names(pred_model_b) <- paste0(names(pred_model_b), "_wo_intervention")
  pred_model <- bind_cols(df_count_psy, pred_model, pred_model_b)
  if (clearance_period) {
    pred_model$clearance_period <- ifelse(pred_model$time_p < as.Date(cut(start_clear_period, breaks = time_period)) ,
                                          "before_pandemic",
                                          ifelse(pred_model$time_p > as.Date(cut(end_clear_period, breaks = time_period)),
                                                 "during_pandemic",
                                                 "during_clearance")
    )
    pred_model$lwr <- ifelse(pred_model$clearance_period == "during_clearance", NA, pred_model$lwr)
    pred_model$upr <- ifelse(pred_model$clearance_period == "during_clearance", NA, pred_model$upr)
    pred_model$fit_wo_intervention <- ifelse(pred_model$time_p >= max(pred_model$time_p[pred_model$clearance_period == "before pandemic"]),
                                             pred_model$fit_wo_intervention,
                                             NA)
  } else {
    pred_model$clearance_period <- pred_model$period
  }
  plot <- ggplot(pred_model, aes_string(x = "time_p", y = count_col)) +
    geom_point(aes(colour = clearance_period)) +
    geom_line(aes(x = time_p, y = fit_wo_intervention, colour = "Expected mean w/o pandemic"), linetype = "dashed") +
    geom_line(aes(x = time_p, y = fit, colour = clearance_period)) +
    geom_ribbon(aes(ymin = lwr, ymax = upr, colour = clearance_period), alpha = 0.3, colour = NA) +
    scale_color_manual(name = "Time Series projection", values = c("before_pandemic" = "black",
                                                                   "Expected mean w/o pandemic" = "red",
                                                                   "during_pandemic" = "black",
                                                                   "during_clearance" = "white")) +
    annotate("text", x = as.Date("2019-10-01"), y = max(pred_model[[count_col]]) + 5, label = annotation) +
    theme(legend.position = "bottom")

@@

```{r}
# Time series analysis of percentage of psy codes per encounter, linear (without seasonality)
patient_count_monthly_psy <- count_icd_psy %>%
  filter(time_p > as.Date(cut( start_date_plots, breaks = time_period)) &
           time_p < as.Date(cut( end_date_plots, breaks = time_period))) %>%
  mutate(time = seq_along(time_p),
         month_year = format(time_p, "%m")
  )

start_clear_period <- as.Date("2020-03-15")
end_clear_period <- as.Date("2020-05-15")
patient_count_monthly_psy_clear <- patient_count_monthly_psy %>%
  mutate( count_psy = ifelse(time_p < as.Date(cut( start_clear_period, breaks = time_period)) |
                               time_p > as.Date(cut( end_clear_period, breaks = time_period)),
                             count_psy,
                             NA)
  )
```

```{r}
ts_linear_count <- lm(count_psy ~ period * time, patient_count_monthly_psy)
ts_linear_count_clear <- lm(count_psy ~ period * time, patient_count_monthly_psy_clear)
ts_seasonal_count <- lm(count_psy ~ period*time + harmonic(month_year, 1, 12), patient_count_monthly_psy)
ts_seasonal_count_clear <- lm(count_psy ~ period*time + harmonic(month_year, 1, 12), patient_count_monthly_psy_clear)
```

### ITS: no clearance period, no seasonality effect

```{r}
time_series_analysis_plotting(ts_linear_count,
                              patient_count_monthly_psy, "count_psy",
                              clearance_period = FALSE) +
  labs(title = "Interrupted Time-Series Analysis: Count of encounters with psychiatric ICD codes",
       subtitle = "Without seasonality effect",
       x = time_period,
       y = "Count")
```

### ITS: clearance period, no seasonality effect

```{r}
time_series_analysis_plotting(ts_linear_count_clear,
                              patient_count_monthly_psy, "count_psy",
                              clearance_period = TRUE) +
  labs(title = "Interrupted Time-Series Analysis: Percentage of encounters with psychiatric ICD codes",
       subtitle = "Without seasonality effect",
       x = time_period,
       y = "Count")
```

### ITS: no clearance period, with seasonality effect

```{r}
time_series_analysis_plotting(ts_seasonal_count,
                              patient_count_monthly_psy, "count_psy",
                              clearance_period = FALSE) +
  labs(title = "Interrupted Time-Series Analysis: Percentage of encounters with psychiatric ICD codes",
       subtitle = "With seasonality effect",
       x = time_period,
       y = "Count")
```

### ITS: with clearance period, with seasonality effect

```{r}
time_series_analysis_plotting(ts_seasonal_count_clear,
                              patient_count_monthly_psy, "count_psy",
                              clearance_period = TRUE) +
  labs(title = "Interrupted Time-Series Analysis: Count of encounters with psychiatric ICD codes",
       subtitle = "With seasonality effect",
       x = time_period,
       y = "Count")
```

>>>>

### Ratio: patients with psychiatric condition / patients without psychiatric condition

@@

ratios_patients_with_without_psy <- patients_without_psy_conditions_counts %>%
  full_join( patients_with_psy_conditions_counts, by=c("time_p", "period")) %>%
  mutate( ratio = count_psy_patients/count_no_psy_patients, 
          percentage = 100 * count_psy_patients/(count_no_psy_patients + count_psy_patients))

```{r}
ratios_patients_with_without_psy %>%
  pivot_longer(cols = c("count_no_psy_patients", "count_psy_patients"), 
               values_to = "count", 
               names_to = "count_type") %>% 
  ggplot(aes(x = time_p, y = count, color = count_type)) +
  geom_point() +
  geom_vline(xintercept = as.Date(pandemic_start_date),
             linetype = "dashed") +
  scale_fill_manual(values = cbPalette) +
  scale_color_manual(values = cbPalette) +
  labs(y = "Count",
       x = paste0("Date (by ", time_period,")"),
       title = paste0("Count comparison: patients with vs without psychiatric conditions (per ", time_period,")")) +
  guides(color = guide_legend(title="Patient diagnoses types"))
```

```{r}
ratio_psy_non_psy <- ratios_patients_with_without_psy %>%
  ggplot(aes(x = time_p, y = ratio, fill = period, color = period)) +
  geom_point() +
  geom_vline(xintercept = as.Date(pandemic_start_date),
             linetype = "dashed") +
  scale_fill_manual(values = cbPalette) +
  scale_color_manual(values = cbPalette) +
  labs(y = "Ratio (with psy conditions/without psy conditions)",
       x = paste0("Date (by ", time_period,")"),
       title = paste0("Ratio patient with vs without psychiatric conditions ( per ", time_period,")"))

@@

```{r}
ratios_patients_with_without_psy %>%
  ggplot(aes(x = time_p, y = percentage, color = period)) +
  geom_point() +
  geom_vline(xintercept = as.Date(pandemic_start_date),
             linetype = "dashed") +
  scale_fill_manual(values = cbPalette) +
  scale_color_manual(values = cbPalette) +
  labs(y = "Percentage",
       x = paste0("Date (by ", time_period,")"),
       title = paste0("Percentage patient with vs total population ( per ", time_period,")")) +
  guides(color = guide_legend(title="Patient diagnoses types"))
```

@@

```{r}
# Model without seasonality
time_series_analysis_plotting <- function(model, df_count_psy, count_col, clearance_period = FALSE) {
  output_model <- broom::tidy(model)
  coeff_slope <- round(output_model[output_model$term == "periodduring_pandemic:time", "estimate"][[1]], 3)
  pvalue_slope <- round(output_model[output_model$term == "periodduring_pandemic:time", "p.value"][[1]], 3)
  annotation <- paste0("Lockdown slope: ", coeff_slope, " (pvalue: ", pvalue_slope, ")")
  pred_model <- predict(model, df_count_psy, type = "response", interval = "confidence") %>% as.data.frame()
  pred_model_b <- predict(model, mutate(df_count_psy, period = "before_pandemic"), type = "response", interval = "confidence") %>%
    as.data.frame()
  names(pred_model_b) <- paste0(names(pred_model_b), "_wo_intervention")
  pred_model <- bind_cols(df_count_psy, pred_model, pred_model_b)
  if (clearance_period) {
    pred_model$clearance_period <- ifelse(pred_model$time_p < as.Date(cut(start_clear_period, breaks = time_period)) ,
                                          "before_pandemic",
                                          ifelse(pred_model$time_p > as.Date(cut(end_clear_period, breaks = time_period)),
                                                 "during_pandemic",
                                                 "during_clearance")
    )
    pred_model$lwr <- ifelse(pred_model$clearance_period == "during_clearance", NA, pred_model$lwr)
    pred_model$upr <- ifelse(pred_model$clearance_period == "during_clearance", NA, pred_model$upr)
    pred_model$fit_wo_intervention <- ifelse(pred_model$time_p >= max(pred_model$time_p[pred_model$clearance_period == "before pandemic"]),
                                             pred_model$fit_wo_intervention,
                                             NA)
  } else {
    pred_model$clearance_period <- pred_model$period
  }
  plot <- ggplot(pred_model, aes_string(x = "time_p", y = count_col)) +
    geom_point(aes(colour = clearance_period)) +
    geom_line(aes(x = time_p, y = fit_wo_intervention, colour = "Expected mean w/o pandemic"), linetype = "dashed") +
    geom_line(aes(x = time_p, y = fit, colour = clearance_period)) +
    geom_ribbon(aes(ymin = lwr, ymax = upr, colour = clearance_period), alpha = 0.3, colour = NA) +
    scale_color_manual(name = "Time Series projection", values = c("before_pandemic" = "black",
                                                                   "Expected mean w/o pandemic" = "red",
                                                                   "during_pandemic" = "black",
                                                                   "during_clearance" = "white")) +
    annotate("text", x = as.Date("2019-10-01"), y = max(pred_model[[count_col]]) + 5, label = annotation) +
    theme(legend.position = "bottom")

@@

```{r}
# Time series analysis of percentage of psy codes per encounter, linear (without seasonality)
patient_count_psy_period_psy <- ratios_patients_with_without_psy %>%
  filter(time_p > as.Date(cut( start_date_plots, breaks = time_period)) &
           time_p < as.Date(cut( end_date_plots, breaks = time_period))) %>%
  mutate(time = seq_along(time_p),
         month_year = format(time_p, "%m")
  )

start_clear_period <- as.Date("2020-03-15")
end_clear_period <- as.Date("2020-05-15")
patient_count_psy_period_psy_clear <- patient_count_psy_period_psy %>%
  mutate( count_no_psy_patients = ifelse(time_p < as.Date(cut( start_clear_period, breaks = time_period)) |
                               time_p > as.Date(cut( end_clear_period, breaks = time_period)),
                             count_no_psy_patients,
                             NA), 
          count_no_psy_patients = ifelse(time_p < as.Date(cut( start_clear_period, breaks = time_period)) |
                               time_p > as.Date(cut( end_clear_period, breaks = time_period)),
                             count_no_psy_patients,
                             NA), 
          ratio = ifelse(time_p < as.Date(cut( start_clear_period, breaks = time_period)) |
                           time_p > as.Date(cut( end_clear_period, breaks = time_period)),
                         ratio,
                         NA),
          percentage = ifelse(time_p < as.Date(cut( start_clear_period, breaks = time_period)) |
                           time_p > as.Date(cut( end_clear_period, breaks = time_period)),
                         percentage,
                         NA)
  )
```

```{r, its count based}
ts_linear_count <- lm(count_psy_patients ~ period * time, patient_count_psy_period_psy)
ts_linear_count_clear <- lm(count_psy_patients ~ period * time, patient_count_psy_period_psy_clear)
ts_seasonal_count <- lm(count_psy_patients ~ period*time + harmonic(month_year, 1, 12), patient_count_psy_period_psy)
ts_seasonal_count_clear <- lm(count_psy_patients ~ period*time + harmonic(month_year, 1, 12), patient_count_psy_period_psy_clear)
```

```{r, its percentage based}
ts_linear_percent <- lm(percentage ~ period * time, patient_count_psy_period_psy)
ts_linear_percent_clear <- lm(percentage ~ period * time, patient_count_psy_period_psy_clear)
ts_seasonal_percent <- lm(percentage ~ period*time + harmonic(month_year, 1, 12), patient_count_psy_period_psy)
ts_seasonal_percent_clear <- lm(percentage ~ period*time + harmonic(month_year, 1, 12), patient_count_psy_period_psy_clear)
```

### ITS: patient counts, no clearance period, no seasonality effect

```{r}
time_series_analysis_plotting(ts_linear_count,
                              patient_count_psy_period_psy,
                              "count_psy_patients",
                              clearance_period = FALSE) +
  labs(title = "Interrupted Time-Series Analysis: Count of encounters with psychiatric ICD codes",
       subtitle = "Without seasonality effect",
       x = time_period,
       y = "Count")
```

### ITS: patient counts, clearance period, no seasonality effect

```{r}
time_series_analysis_plotting(ts_linear_count_clear,
                              patient_count_psy_period_psy,
                              "count_psy_patients",
                              clearance_period = TRUE) +
  labs(title = "Interrupted Time-Series Analysis: Percentage of encounters with psychiatric ICD codes",
       subtitle = "Without seasonality effect",
       x = time_period,
       y = "Count")
```

### ITS: patient counts, no clearance period, with seasonality effect

```{r}
time_series_analysis_plotting(ts_seasonal_count,
                              patient_count_psy_period_psy, "count_psy",
                              clearance_period = FALSE) +
  labs(title = "Interrupted Time-Series Analysis: Percentage of encounters with psychiatric ICD codes",
       subtitle = "With seasonality effect",
       x = time_period,
       y = "Count")
```

### ITS: patient counts, with clearance period, with seasonality effect

```{r}
time_series_analysis_plotting(ts_seasonal_count_clear,
                              patient_count_psy_period_psy, "count_psy",
                              clearance_period = TRUE) +
  labs(title = "Interrupted Time-Series Analysis: Count of encounters with psychiatric ICD codes",
       subtitle = "With seasonality effect",
       x = time_period,
       y = "Count")
```

### ITS: psy percentage, no clearance period, no seasonality effect

```{r}
time_series_analysis_plotting(ts_linear_percent,
                              patient_count_psy_period_psy,
                              "percentage",
                              clearance_period = FALSE) +
  labs(title = "Interrupted Time-Series Analysis: Percentage of encounters with psychiatric ICD codes",
       subtitle = "Without seasonality effect",
       x = time_period,
       y = "Percentage")
```

### ITS: psy percentage, clearance period, no seasonality effect

```{r}
time_series_analysis_plotting(ts_linear_percent_clear,
                              patient_count_psy_period_psy,
                              "percentage",
                              clearance_period = TRUE) +
  labs(title = "Interrupted Time-Series Analysis: Percentage of encounters with psychiatric ICD codes",
       subtitle = "Without seasonality effect",
       x = time_period,
       y = "Percentage")
```

### ITS: psy percentage, no clearance period, with seasonality effect

```{r}
time_series_analysis_plotting(ts_seasonal_percent,
                              patient_count_psy_period_psy,
                              "percentage",
                              clearance_period = FALSE) +
  labs(title = "Interrupted Time-Series Analysis: Percentage of encounters with psychiatric ICD codes",
       subtitle = "Without seasonality effect",
       x = time_period,
       y = "Percentage")
```

### ITS: psy percentage, with clearance period, with seasonality effect

```{r}
time_series_analysis_plotting(ts_seasonal_percent_clear,
                              patient_count_psy_period_psy,
                              "percentage",
                              clearance_period = TRUE) +
  labs(title = "Interrupted Time-Series Analysis: Percentage of encounters with psychiatric ICD codes",
       subtitle = "With seasonality effect",
       x = time_period,
       y = "Percentage")
```
